Future challenges for Research Software Engineering

### 2025-05-18 future_challenges_for_RSEng.tex
Update future_challenges_for_RSEng.tex
Co-authored-by: Gerasimos Chourdakis <gerasimos.chourdakis@ipvs.uni-stuttgart.de>

\subsection{Impact of AI and automation}
\begin{whatis}{}{impact-ai-automation}
RSEs understand how AI models work, and can help researchers reap their benefits, while
clearly knowing their limits.
\end{whatis}
Generative AI tools, such as GitHub Copilot \cite{Friedmann2021}, are being added to software development tools.
These AI powered code assistants promise to increase productivity by automatically generating code, and writing test cases and documentation.
Software developers use them to reduce the number of keystrokes, complete tasks and to recall syntax without having to consult the documentation \cite{Liang2024}.
Code assistants have the potential to help novice or even non-expert programmers write code \cite{Feldman2024} thus enabling non-experts to participate in research and software creation, and fostering greater collaboration across disciplines.
However, when these tools are used in practice many users report increased cognitive load and frustration stemming from the difficulty to prompt the system effectively and debugging the generated code \cite{Simkute2025}.

>>>>

\subsection{Impact of AI and automation}
\begin{whatis}{}{impact-ai-automation}
RSEs understand how AI models work, and can help researchers reap their benefits, while
clearly knowing their limits.
\end{whatis}
Generative AI tools, such as GitHub Copilot \cite{Friedmann2021}, are being added to software development tools.
These AI powered code assistants promise to increase productivity by automatically generating code, and writing test cases and documentation.
Software developers use them to reduce the number of keystrokes, complete tasks and to recall syntax without having to consult the documentation \cite{Liang2024}.
Code assistants have the potential to help novice or even non-expert programmers write code \cite{Feldman2024} thus enabling non-experts to participate in research and software creation, and fostering greater collaboration across disciplines.
However, when these tools are used in practice, many users report increased cognitive load and frustration stemming from the difficulty to prompt the system effectively and from debugging the generated code \cite{Simkute2025}.
